Canopy leaf area of a mature evergreen *Eucalyptus* woodland does not respond to elevated atmospheric [CO~2~] but tracks water availability

### 2015-08-15 manuscript.Rmd
Litter methods moved from results; some notes added for revision

@@

\
An outbreak of psyllids (*Cardiaspina* sp.), a tiny sap-feeding insect, resulted in rapid loss of canopy leaf area in one of the rings (ring 6), starting in September 2014 [for description of a similar outbreak in *Eucalyptus moluccana*, see @Hall2015]. This rapid decrease in LAI was evident in a quick increase in $\tau_d$ (Fig. 1). However, it recovered rapidly after rainfall in December 2014, resulting in a lower $\tau_d$ than before the insect attack.

\
We found no differences in the LAI anomaly between the C~a~ treatments, as shown by consistent overlap of the 95% confidence intervals of the generalized additive model  (thus, p > 0.05 at any point in time, Fig. 3). A brief exception was the period where one of the rings was affected by an insect outbreak (see above). A linear mixed-effects model with C~a~ and time as fixed effects, and FACE ring as a random effect also showed no significant effect of C~a~ on the LAI anomaly (p `r pval(LanomAOV[[3]][1])`). But what difference between C~a~ treatments could we have reasonably detected? The average half-width of the confidence interval on the LAI anomaly (Fig. 4) was `r f(LanomCIamb)` m^2^m^-2^ for aC~a~, and `r f(LanomCIele)` m^2^m^-2^ for the eC~a~ treatment. This means that, provided the within-treatment variance stays the same, we could detect a difference of `r f(LanomCIamb + LanomCIele)` m^2^ m^-2^ (the sum of the half-width CI's), or about a change of `r f(Ldetect)`% over the mean LAI of the EucFACE (`r f(mean(ba$LAI.mean))` m^2^m^-2^).

\
The dynamics of $LAI_\tau$ were compared with litter fall, by estimating the absolute change in $LAI_\tau$ for every FACE ring, for every period when litter was collected. We did this by fitting a generalized additive model to $LAI_\tau$ for each ring separately (similar to that shown for $\tau_d$ in Fig. 1), and calculating the change in $LAI_\tau$ for each period ($\Delta L_\tau$). Litter production correlated significantly with $\Delta L_\tau$ when the latter was separated into periods where $\Delta L_\tau$ > 0 (net leaf growth) and $\Delta L_\tau$ <  0 (net leaf loss) (Fig. 4). During periods of net leaf loss, we found a significant correlation (with a slope of `r f(tidy(lmlitdown)$estimate[2])`, SE = `r f(tidy(lmlitdown)$std.error[2])`, n = `r glance(lmlitdown)$df.residual+2`), close to the 1:1 line, as expected since no leaf growth was observed during any period where LAI was decreasing. During periods of net leaf growth, we found a significant correlation (with a slope of `r f(tidy(lmlitup)$estimate[2])`, SE = `r f(tidy(lmlitup)$std.error[2])`, n = `r glance(lmlitup)$df.residual+2`) between $\Delta L_\tau$ and litter production (Fig. 5).

\
We estimated gross leaf production from measured litter production (P~L~) and temporal dynamics of $LAI_\tau$. For each litter collection period, the net change in $LAI_\tau$ was estimated using fitted generalized additive models to allow interpolation from a smoothed timeseries. Gross leaf production was then calculated as the sum of the net change in $LAI_\tau$ and the amount of leaf litter produced during that period. The temporal dynamics of litter and leaf production are shown in Fig. 5, and demonstrate the irregular temporal pattern in leaf production and P~L~. We found no significant difference in either litter or leaf production between C~a~ treatments at any time point (p > 0.2 for both, assessed with linear-mixed effects models).

>>>>

\ To compare the dynamics of LAI with litter production, we calculated the change in LAI for every time period when litter was collected. We did this by fitting a generalized additive model to LAI for each ring separately (similar to that shown for $\tau_d$ in Fig. 1), and calculating the change in LAI for each period ($\Delta L_\tau$). Next, we estimated gross leaf production from measured litter production (P~L~) and temporal dynamics of $LAI_\tau$. For each litter collection period, the net change in $LAI_\tau$ was estimated using fitted generalized additive models to allow interpolation from a smoothed timeseries. Gross leaf production was then calculated as the sum of the net change in $LAI_\tau$ and the amount of leaf litter produced during that period.

@@

\
An outbreak of psyllids (*Cardiaspina* sp.), a tiny sap-feeding insect, resulted in rapid loss of canopy leaf area in one of the rings (ring 6), starting in September 2014 [for description of a similar outbreak in *Eucalyptus moluccana*, see @Hall2015; pers. comm. Andrew Gherlenda, Markus Riegler]. This rapid decrease in LAI was evident in a quick increase in $\tau_d$ (Fig. 1). However, it recovered rapidly after rainfall in December 2014, resulting in a lower $\tau_d$ than before the insect attack.

\
We found no differences in LAI between the C~a~ treatments, as shown by consistent overlap of the 95% confidence intervals of the generalized additive model (thus, p > 0.05 at any point in time, Fig. 3). A brief exception was the period where one of the rings was affected by an insect outbreak (see above). A linear mixed-effects model with C~a~, time and pre-treatment basal area (Fig. 3) as fixed effects, and FACE ring as a random effect also showed no significant effect of C~a~ on LAI (p **fix this**). 
** rethink the following; CI on least-square means accounting for BA?**
But what difference between C~a~ treatments could we have reasonably detected? The average half-width of the confidence interval on the LAI anomaly (Fig. 4) was `r f(LanomCIamb)` m^2^m^-2^ for aC~a~, and `r f(LanomCIele)` m^2^m^-2^ for the eC~a~ treatment. This means that, provided the within-treatment variance stays the same, we could detect a difference of `r f(LanomCIamb + LanomCIele)` m^2^ m^-2^ (the sum of the half-width CI's), or about a change of `r f(Ldetect)`% over the mean LAI of the EucFACE (`r f(mean(ba$LAI.mean))` m^2^m^-2^).

\
Litter production correlated significantly with $\Delta L_\tau$ when the latter was separated into periods where $\Delta L_\tau$ > 0 (net leaf growth) and $\Delta L_\tau$ <  0 (net leaf loss) (Fig. 4). During periods of net leaf loss, we found a significant correlation (with a slope of `r f(tidy(lmlitdown)$estimate[2])`, SE = `r f(tidy(lmlitdown)$std.error[2])`, n = `r glance(lmlitdown)$df.residual+2`), close to the 1:1 line, as expected since no leaf growth was observed during any period where LAI was decreasing. During periods of net leaf growth, we found a significant correlation (with a slope of `r f(tidy(lmlitup)$estimate[2])`, SE = `r f(tidy(lmlitup)$std.error[2])`, n = `r glance(lmlitup)$df.residual+2`) between $\Delta L_\tau$ and litter production (Fig. 5).

\
The temporal dynamics of litter and leaf production are shown in Fig. 5, and demonstrate the irregular temporal pattern in leaf production and P~L~. We found no significant difference in either litter or leaf production between C~a~ treatments at any time point (p > 0.2 for both, assessed with linear-mixed effects models).
